---
title: "Exploratory Data Analysis - Retail"
author: "Aditya Dawn"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Objective**  
Here, we are given to do Exploratory Data Analysis on a Super Store.  

#### *Tool Used*  
* R and Python for Data Analysis.  
* R Markdown for this report.    

##### Set-up of R Markdown  
Before starting Analysis, I want to make my work-space suitable for the use of both of R and Python.  
```{r echo=TRUE}
library(reticulate)
use_python("/usr/local/bin/python")
use_virtualenv("~/myenv")
use_condaenv("myenv")
```

Now we are ready to do the Data Analysis.  

## **Import of Data**  
```{r echo=TRUE}
df <- read.csv("/home/adi/Documents/Aditya/TSF-GRIPFEB21/SampleSuperstore.csv", header = TRUE)
```  
## **Pre-processing of the Data**  
For, making the Data suitable for Analysis, the $DataExplorer$ of R package is used.  
```{r echo=TRUE}
library(DataExplorer)
```

### *Missing Value Treatment*  
```{r echo=TRUE}
plot_missing(df)
```
The plot clearly shows that there are no missing values.  

### *Variables*  
```{r echo=TRUE}
library(dplyr)
```

* *Categorical Variables*  

```{r echo=TRUE}
df %>% select_if(is.character) %>% head(2)
```
The above output shows the features with class "Character". In this case, we have 8 Categorical Features.  

* *Numerical Variables*  
```{r echo=TRUE}
df %>% select_if(is.numeric) %>% head(2)
```
The above output shows the features with class "Numeric". In this case, we have 5 feature with class "Numeric".  

#### *Removal of unwanted features*  
The "Postal.Code" feature will not be helping much in this context. So, we can remove it.  
```{r echo=TRUE}
df <- df[,-6]
```



```{python echo=TRUE}
import matplotlib.pyplot as plt
import seaborn as sns
```

## **Visualization of Categorical Variables**  
1. *Ship Mode*  

```{python echo=TRUE}
r.df['Ship.Mode'].unique()
```
```{python echo=TRUE}
sns.countplot(x=r.df['Ship.Mode'])
plt.show()
```


2. *Segment*  

```{python echo=TRUE}
r.df['Segment'].unique()
```

```{python echo=TRUE}
sns.countplot( x = r.df['Segment'] )
plt.show()
```


3. *Country*  
```{python echo=TRUE}
r.df['Country'].unique()
```
So, the feature "Country" has only one character. So, we can remove it.  
```{r echo=TRUE}
df <- df[,-3]
```

4. *City*  
```{python echo=TRUE}
r.df['City'].unique()
```

5. *State*  

```{python echo=TRUE}
r.df['State'].unique()
```
```{python echo=TRUE}
plt.figure(figsize = (10,10))
sns.countplot(x=r.df['State'])
plt.xticks(rotation = -90)
plt.show()
```

6. *Region*  

```{python echo=TRUE}
r.df['Region'].unique()
```
```{python echo=TRUE}
sns.countplot(x=r.df['Region'])
plt.show()
```

7. *Category*  
```{python echo=TRUE}
r.df['Category'].unique()
```
```{python echo=TRUE}
sns.countplot(x=r.df['Category'])
plt.show()
```

8. *Sub-Category*  
```{python echo=TRUE}
r.df['Sub.Category'].unique()
```
```{python echo=TRUE}
sns.countplot(x=r.df['Sub.Category'])
plt.xticks(rotation = -45)
plt.show()
```

## **Visualization of Numerical Variables**  

```{python echo=TRUE}
r.df.hist(bins=50 ,figsize=(20,15))
plt.show()
```


```{python echo=TRUE}
sns.pairplot(r.df,hue='Sub.Category')
plt.show()
```



```{python echo=TRUE}
sns.heatmap(r.df.corr())
```

### **Analysis of the above visualizations taking two at a time**  
1. *Sales vs Quantity*  
```{r echo=TRUE}
library(ggplot2)
ggplot(data = df, 
       aes(x = Quantity, y = Sales, fill = Ship.Mode) ) +
  geom_bar(stat = "identity") 
```
**Clearly, most of the sales have been caused by the standard class of shipment mode.**  

2. *Profit vs Sales* (based on Ship Mode)  
```{r echo=TRUE}
ggplot(data = df, 
       aes(x = Sales, y = Profit, color = Ship.Mode)) + 
  geom_point()
```
**As a result, more profit and loss have been availed from the standard shipment class. But, there are not higher range profits seen due to the other features.**  

3. *Sales vs Discount*  
```{r echo=TRUE}
ggplot() + geom_point(data = df, 
                      aes(x = Discount, 
                          y = Sales, 
                          color = Ship.Mode)) 
```

* **From the above graph, it can be seen that sales are attracted more by discounts.**  
* **Discounts attract mostly the Standard Class shipment.**  
* **Same day shipment mode get the least discount offers.**  

4. *Profit vs Discount*  

```{r echo=TRUE}
ggplot() + 
  geom_bar(data = df, 
           aes(x = Discount, y = Profit, fill = Ship.Mode), stat = "identity") 
```
**Here it is clear that, Discount and Profit has a kind of inverse relation with each other. The segments with more discounts, made lesser profits.**  

5. *Profit vs Sub-Category*  
```{r echo=TRUE}
ggplot() + 
  geom_bar(data = df, aes(x = Sub.Category, y = Profit, fill = Region), stat = "identity") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
**From the above diagram, it is clear that the Blinders faced maximum loss, followed by the Tables and the Machines in the Central Region.**  

6. *Sales vs Category*  
```{r echo=TRUE}
ggplot() + 
  geom_bar(data = df, aes(x = Category, y = Sales, fill = Region), stat = "identity") +       
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
**The above diagram shows that, Technology made most of the sales, followed by Furniture and Office Supplies, mainly in the regions of West, followed by Central.**  

7. *Profit vs Category*  

```{r echo=TRUE}
ggplot() + 
  geom_bar(data = df, aes(x = Category, y = Profit, fill = Region), stat = "identity") +      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
**Here it can be seen that, the maximum profit is made by Technology, followed by Office Supplies and Furniture.**  

8. *Profit vs Sales* (based on Category)  
```{r echo=TRUE}
ggplot() + geom_point(data = df, aes(x = Sales, y = Profit, color = Category)) 
```
**Here, it is clear that, the relation between Profit and Sale is more of less linear.**  



## **Insights**   
* Shipment Mode other than Standard Class should be promoted more.  
* Unnecessary Discounts with low rates can face risk.  
* The loss in Furniture should be taken care.  
* It is requested to look into the sales growth of the Categories in the Regions of East and South.  
  
  
## THANK YOU!  
